---
title: "Downstream_Analyses"
author: "Sushant Patkar"
date: "2024-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Define Paths to clinical metadata, HistoTME prediction files, IHC Validation files, output directory
```{r}
CLINICAL_METADATA_PATH = "~/Tamara_ICB_response/SUNY_FINAL_clinical_data_with_PFS_corrected_050224.tsv"
HISTOTME_SUNY_NSCLC = "~/Tamara_ICB_response/suny_predictions_multitask_UNI.csv"
HISTOTME_CPTAC_NSCLC = "~/Tamara_ICB_response/cptac_predictions_multitask_UNI.csv"
HISTOTME_TCGA_NSCLC = "~/Tamara_ICB_response/TCGA_predictions_multitask_UNI.csv"
CD8_QUPATH_MEASUREMENTS = "~/SUNY_CD4_CD8/SUNY_cd8_tcell_qupath_measurements.csv"
CD4_QUPATH_MEASUREMENTS = "~/SUNY_CD4_CD8/SUNY_cd4_tcell_qupath_measurements.csv"
BCELL_QUPATH_MEASUREMENTS = "~/SUNY_CD3_CD20/SUNY_bcell_qupath_measurements_v2.csv"
MAC_QUPATH_MEASUREMENTS = "~/SUNY_P40_CD163/SUNY_mac_qupath_measurements.csv"
OUTPUT_DIR = "~/Tamara_ICB_response"
```

# Load required R libraries and packages
```{r}
library(readxl)
library(survival)
library(survminer)
library(caret)
library(ConsensusClusterPlus)
library(ComplexHeatmap)
library(factoextra)
library(cluster)
library(ggplot2)
library(pROC)
library(epiR)
```

# Read clinical metadata and HistoTME signature predictions
```{r}
clindat <- read_delim(CLINICAL_METADATA_PATH,delim = "\t")
clindat <- clindat[!duplicated(clindat$case_id),]
suny_predictions <- read_csv(HISTOTME_SUNY_NSCLC)
tcga_predictions <- read_csv(HISTOTME_TCGA_NSCLC)
cptac_predictions <- read_csv(HISTOTME_CPTAC_NSCLC)
```

# Unsupervised Analysis: Perform k-means clustering of TCGA+CPTAC data and build classification model based on cluster labels
```{r}
# merge HistoTME predictions from TCGA and CPTAC datasets
cptac_predictions <- as.data.frame(cptac_predictions)
cptac_exp <- cptac_predictions[,2:(ncol(cptac_predictions))]
cptac_exp <- apply(as.matrix(cptac_exp),2,scale)
rownames(cptac_exp) <- cptac_predictions$ID
tcga_predictions <- as.data.frame(tcga_predictions)
tcga_exp <- tcga_predictions[,2:(ncol(tcga_predictions))]
tcga_exp <- apply(as.matrix(tcga_exp),2,scale)
rownames(tcga_exp) <- tcga_predictions$ID
mat <- rbind(tcga_exp,cptac_exp)

# evaluate K means clustering metrics for different choices of K
res = ConsensusClusterPlus(d = t(mat), distance = "euclidean", clusterAlg = "km", pItem = 0.8, pFeature = 1, seed = 42,maxK = 6, reps = 100)
fviz_nbclust(mat, kmeans, method = "silhouette")
pdf(paste(OUTPUT_DIR,"Kmeans_Silhouette_scores.pdf",sep = "/"))
fviz_nbclust(mat, kmeans, method = "silhouette")
dev.off()


# Perform PCA
pca_result <- prcomp(mat, scale. = TRUE)
pc_scores <- as.data.frame(pca_result$x)
pc_scores$Cluster <- (c("Immune-Desert","Immune-Inflammed")[as.numeric(as.character(res[[2]]$consensusClass) == "2")+1])

# Plot PCA using ggplot2
# Create ggplot object for 3D PCA
pca_plot <- ggplot(pc_scores, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(title = "PCA",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  scale_color_manual(values = c( "#88CCEE","#882255")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

print(pca_plot)

# Train a random-forest classifier to predict cluster membership
set.seed(42)
control <- trainControl(method = "cv", number=5,)
df <- data.frame((mat), clus = factor(as.character(res[[2]]$consensusClass)))
model <- suppressWarnings(train(clus ~ ., data = df, "rf", trainControl = control))

# Plot feature importances
fimp <- varImp(model)$importance
fimp_plot <- ggplot(data = data.frame(Feature = factor(rownames(fimp), levels = rownames(fimp)[order(fimp$Overall, decreasing = T)]), Importance = fimp$Overall), aes(x = Feature, y = Importance)) + geom_bar(stat = "identity") + theme_minimal() + theme(text = element_text(angle = 90, size = 15)) + xlab("")
print(fimp_plot)
ggsave(plot = fimp_plot, filename = paste(OUTPUT_DIR,"fimp_plot.pdf",sep = "/"), width = 10, height = 6)
```

# Classify unseen samples from SUNY based on TME classification model
```{r}
suny_predictions <- as.data.frame(suny_predictions)
suny_exp <- suny_predictions[,2:(ncol(suny_predictions))]
suny_exp <- apply(as.matrix(suny_exp),2,scale)
rownames(suny_exp) <- suny_predictions$ID
colnames(suny_exp) <- colnames(suny_predictions)[2:(ncol(suny_predictions))]
preds <- predict(model, suny_exp)
```

# Merge HistoTME predictions and clinical metadata of SUNY patients
```{r}
clindat$CCI[clindat$CCI == "≥8"] = "8"
survdat <- data.frame(OS_time = clindat$OS_time, OS_status = clindat$OS_status, DSS_time = clindat$DSS_time, DSS_status = clindat$DSS_status, score = NA, treatment = clindat$io_cohort, stage = clindat$cAJCC, grade = clindat$dx_ca_grade,main_adj = !is.na(clindat$io_main) | !is.na(clindat$io_adj), firstline = !is.na(clindat$io_1L), subtype = clindat$histological_type_at_dx, smoking_hx = clindat$smoking_hx, CCI = scale(as.numeric(clindat$CCI)), age = (as.numeric(clindat$age_at_dx)), sample_type = c("biopsy","resection")[as.numeric(grepl("UR-PDL1-LR",clindat$case_id))+1], PDL1_scores = clindat$PDL1_scores, NL_ratio = scale(as.numeric(clindat$`IO_start_neut-lymph_ratio`)), overall_resp = clindat$overall_resp, PFS_adj_time =  clindat$PFS_adj_time, PFS_adj_status = clindat$PFS_adj_status, PFS_1L_time = clindat$PFS_1L_time, PFS_1L_status = clindat$PFS_1L_status, PFS_2L_time = clindat$PFS_2L_time, PFS_2L_status = clindat$PFS_2L_status, PFS_SL_time = clindat$PFS_SL_time, PFS_SL_status = clindat$PFS_SL_status, io_main = clindat$io_main, io_adj = clindat$io_adj, io_1L = clindat$io_1L, resp_1L = clindat$first_line_resp, tmb = scale(clindat$TMB), io_2L = clindat$io_2L, io_SL = clindat$io_SL, overall_resp_stringent = clindat$overall_resp_stringent, chemo_1L = clindat$chemo_1L, chemo_2L = clindat$chemo_2L, chemo_SL = clindat$chemo_SL, chemo_adj = clindat$chemo_adj, chemo_main = clindat$chemo_main, sex = clindat$sex, race = clindat$race)

survdat$io_main_type <- ""
survdat$io_adj_type <- ""
survdat$io_1L_type <- ""
survdat$io_2L_type <- ""
survdat$io_SL_type <- ""

survdat$io_main_type[!is.na(survdat$io_main) &!is.na(survdat$chemo_main)] <- "io+chemo"
survdat$io_main_type[!is.na(survdat$io_main) & is.na(survdat$chemo_main)] <- "io"
survdat$io_adj_type[!is.na(survdat$io_adj) &!is.na(survdat$chemo_adj)] <- "io+chemo"
survdat$io_adj_type[!is.na(survdat$io_adj) & is.na(survdat$chemo_adj)] <- "io"
survdat$io_1L_type[!is.na(survdat$io_1L) &!is.na(survdat$chemo_1L)] <- "io+chemo"
survdat$io_1L_type[!is.na(survdat$io_1L) & is.na(survdat$chemo_1L)] <- "io"
survdat$io_2L_type[!is.na(survdat$io_2L) &!is.na(survdat$chemo_2L)] <- "io+chemo"
survdat$io_2L_type[!is.na(survdat$io_2L) & is.na(survdat$chemo_2L)] <- "io"
survdat$io_SL_type[!is.na(survdat$io_SL) &!is.na(survdat$chemo_SL)] <- "io+chemo"
survdat$io_SL_type[!is.na(survdat$io_SL) & is.na(survdat$chemo_SL)] <- "io"

site <- clindat$prim_met_res_site
site[is.na(site)] <- clindat$prim_met_bx1_site[is.na(site)]
site[is.na(site)] <- clindat$prim_met_bx2_site[is.na(site)]
survdat$site <- site
survdat$subtype[!(survdat$subtype %in% c("NSCLC_adenoca","NSCLC_sqcc")) & !is.na(survdat$subtype)] <- "other"
rownames(survdat) <- clindat$case_id
survdat <- survdat[rownames(suny_exp),]


survdat$cluster <- preds
survdat$response_group <- strata(c("PD-L1 < 50%","PD-L1 ≥50%")[as.numeric(survdat$PDL1_scores %in% c("≥50%"))+1], c("Immune Desert","Immune Inflamed")[as.numeric(survdat$cluster == "2")+1])
```

# Plot HistoTME predictions for the SUNY cohort
```{r}
ha = columnAnnotation(
    site = survdat$site,
    subtype = survdat$subtype,
    grade = survdat$grade,
    stage = survdat$stage,
    specimen_type = survdat$sample_type,
    treatment = c("no io","received io")[as.numeric(survdat$treatment == "received io")+1],
    col = list("site" = c("Primary" = "#009E73", 
                          "Metastatic" =  "#F0E442"),
                          
              "stage" = c("I"="#DF65B0",
                          "II" = "#984EA3",
                          "III" = "#FEE5D9",
                          "IV" = "#FB6A4A"),
               "grade" = c("G1" = "#db6d00", 
                           "G2" = "#924900",
                           "G3" = "#525252",
                           "G4" = "#F2F0F7"),
               "subtype" = c("NSCLC_adenoca" = "#CE1256", "NSCLC_sqcc" =  "#6A51A3", "other" = "#01FFAB"),
              "specimen_type" = c("resection" = "#D95F02", "biopsy" = "#1F78B4"),
              "treatment" = c("received io" = "#F00932","no io" = "#CCFFFF")
    )
)

mat <- suny_exp
rownames(mat) <- NULL
Heatmap(t(mat), row_order = order(varImp(model)$importance[,1], decreasing = T), column_split = preds, cluster_columns= T, cluster_column_slices = FALSE, column_gap = unit(2, "mm"),heatmap_legend_param = list(title = "score"), top_annotation = ha)

#pdf(paste(OUTPUT_DIR,"Heatmap_unsupervised_SUNY2.pdf",sep = "/"), width = 16, height = 5)
#Heatmap(t(mat),  row_order = order(varImp(model)$importance[,1], decreasing = T), column_split = preds, cluster_columns= T, cluster_column_slices = FALSE, column_gap = unit(2, "mm"),heatmap_legend_param = list(title = "score"), top_annotation = ha)
#dev.off()
```
# Validate TME class predictions against matched IHC data of resection cases
```{r}
tcell_qupath_measurements <- read_csv(CD8_QUPATH_MEASUREMENTS)
tcell_qupath_measurements2 <- read_csv(CD4_QUPATH_MEASUREMENTS)
bcell_qupath_measurements <- read_csv(BCELL_QUPATH_MEASUREMENTS)
mac_qupath_measurements <- read_csv(MAC_QUPATH_MEASUREMENTS)

cc <- as.character(preds)
names(cc) <- rownames(suny_exp)
dd_clus_tcell <- data.frame(density = (tcell_qupath_measurements$`Num Positive per mm^2` + tcell_qupath_measurements2$`Num Positive per mm^2`), clus = cc[substr(tcell_qupath_measurements$Image,1,14)])
dd_clus_tcell$Image <- tcell_qupath_measurements$Image
dd_clus_tcell <- dd_clus_tcell[!substr(dd_clus_tcell$Image,1,14) %in% c("UR-PDL1-LR-041","UR-PDL1-LR-048","UR-PDL1-LR-083"),]

p1 <- ggplot(data = dd_clus_tcell[!is.na(dd_clus_tcell$clus),], aes(x = clus, y = density)) + geom_boxplot() + theme_bw() + xlab("H&E Predicted cluster") + ylab("CD4+CD8 IHC (num cells/mm^2)") + stat_compare_means()
print(p1)

dd_clus_bcell <- data.frame(density = (bcell_qupath_measurements$`Num Positive per mm^2`), clus = cc[substr(bcell_qupath_measurements$Image,1,14)])
dd_clus_bcell$Image <- bcell_qupath_measurements$Image
dd_clus_bcell <- dd_clus_bcell[!substr(dd_clus_bcell$Image,1,14) %in% c("UR-PDL1-LR-041","UR-PDL1-LR-048","UR-PDL1-LR-083"),]

p2 <- ggplot(data = dd_clus_bcell[!is.na(dd_clus_bcell$clus),], aes(x = clus, y = density)) + geom_boxplot() + theme_bw() + xlab("H&E Predicted cluster") + ylab("CD3+CD20 IHC (num cells/mm^2)") + stat_compare_means()
print(p2)

dd_clus_mac <- data.frame(density = (mac_qupath_measurements$`Num Positive per mm^2`), clus = cc[substr(mac_qupath_measurements$Image,1,14)])
dd_clus_mac$Image <- mac_qupath_measurements$Image
dd_clus_mac <- dd_clus_mac[!substr(dd_clus_mac$Image,1,14) %in% c("UR-PDL1-LR-041","UR-PDL1-LR-048","UR-PDL1-LR-083"),]

p3 <- ggplot(data = dd_clus_mac[!is.na(dd_clus_mac$clus),], aes(x = clus, y = density)) + geom_boxplot() + theme_bw() + xlab("H&E Predicted cluster") + ylab("CD163+P40 IHC (num cells/mm^2)") + stat_compare_means()
print(p3)


dd_all <- rbind(dd_clus_tcell,rbind(dd_clus_bcell, dd_clus_mac))
dd_all$cell_type <- c(rep("T cell", nrow(dd_clus_tcell)),rep("B cell", nrow(dd_clus_bcell)),rep("Macrophage", nrow(dd_clus_mac)))
dd_all$cell_type <- factor(dd_all$cell_type, levels = c("T cell","B cell","Macrophage"))

p4 <- ggboxplot(dd_all[!is.na(dd_all$clus),], x = "clus", y="density", facet.by = "cell_type") + theme_bw() + theme(text = element_text(size = 15)) + stat_compare_means(comparisons = list(c("2","1")), method = "wilcox", label = "p.signif") + xlab("AI Prediction") + ylab("# positive cells/mm^2")
print(p4)

#ggsave(plot = p4, filename = paste(OUTPUT_DIR,"unsupervised_IHC_validation.pdf"), width = 7, height = 5)

```

# Associations between TME class, PD-L1 scores and overall survival of ICI treated patients
```{r}
# Association between TME class and overall survival 
df <- survdat[!is.na(survdat$io_1L),]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO")
A_1L <- p1


# Association between PD-L1 scores and overall survival 
df <- survdat[!is.na(survdat$io_1L),]
surv <- survfit(Surv(OS_time, OS_status) ~ PDL1_scores, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ PDL1_scores, data = df))



p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
#legend.labs=c('<1%', '1-49%','≥50%'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#E7298A","#7B3294","#885000","#88CCEE","#882255","#CC6677","#7B3254"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO")
B_1L <- p1

# Association between TME class and overall survival in PD-L1 <1% subgroup
df <- survdat[!is.na(survdat$io_1L) & survdat$PDL1_scores %in% c("<1%"),]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO (PDL1 <1%)")
C_1L <- p1


# Association between TME class and overall survival in PD-L1 1-49% subgroup
df <- survdat[!is.na(survdat$io_1L) & survdat$PDL1_scores %in% c("1-49%"),]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO (PDL1 TPS 1-49%)")
D_1L <- p1

# Association between TME class and overall survival in PD-L1 >= 50% subgroup
df <- survdat[!is.na(survdat$io_1L) & survdat$PDL1_scores %in% c("≥50%"),]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO (PDL1 TPS ≥50%)")
E_1L <- p1


#combine all K-M plots into one figure
fig1 <- ggarrange(A_1L$plot,B_1L$plot,A_1L$table, B_1L$table, ncol = 2, nrow = 2,labels = c("A","B"),heights = c(3,1))
fig1_1 <- ggarrange(C_1L$plot, D_1L$plot, E_1L$plot,C_1L$table,D_1L$table,E_1L$table, ncol = 3, nrow = 2, labels = c("C","D","E"), heights = c(3,1))
fig_combined <- ggarrange(fig1,fig1_1, nrow = 2, ncol = 1)
print(fig_combined)

#ggsave(plot = fig_combined,paste(OUTPUT_DIR,"updated_survival_figure_main.pdf",sep = "/"), width = 16, height = 13)


# Association between TME class and overall survival for patients receiving first line ICI+chemo
df <- survdat[!is.na(survdat$io_1L) & survdat$io_1L_type == "io+chemo",]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO + Chemo")
AA1 <- p1

# Association between TME class and overall survival for patients receiving first line ICI monotherapy
df <- survdat[!is.na(survdat$io_1L) & survdat$io_1L_type == "io",]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 20)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "OS First line IO monotherapy")
AA2 <- p1

# combine all K-M plots into one figure
fig1 <- ggarrange(AA1$plot,AA2$plot,AA1$table, AA2$table, ncol = 2, nrow = 2,labels = c("A","B"),heights = c(3,1))
print(fig1)
#ggsave(plot = fig1,paste(OUTPUT_DIR,"survival_figure_treatment_type_1L.pdf",sep = "/"), width = 10, height = 5)


df <- survdat[survdat$treatment == "received io",]
surv <- survfit(Surv(OS_time, OS_status) ~ cluster, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "Overall survival (all IO patients)")
A<- p1


df <- survdat[survdat$treatment == "received io",]
surv <- survfit(Surv(OS_time, OS_status) ~ PDL1_scores, data = df)
summary(coxph(Surv(OS_time, OS_status) ~ PDL1_scores, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
#legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#E7298A","#7B3294","#885000","#88CCEE","#882255","#CC6677","#7B3254"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "Overall survival (all IO patients)")
B <- p1

# combine all K-M plots into one figure
fig1 <- ggarrange(A$plot,B$plot,A$table, B$table, ncol = 2, nrow = 2,labels = c("A","B"),heights = c(3,1))
print(fig1)
#ggsave(plot = fig1,paste(OUTPUT_DIR,"updated_survival_figure2.pdf"), width = 14, height = 8)
```

# Associations between TME class, PD-L1 scores and progression free survival of ICI treated patients
```{r}
# Association between TME class and progression free survival for patients receiving first line ICI 
df <- survdat[!is.na(survdat$io_1L),]
surv <- survfit(Surv(PFS_1L_time, PFS_1L_status) ~ cluster, data = df)
summary(coxph(Surv(PFS_1L_time, PFS_1L_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS First line IO")
A_11L <- p1


# Association between PD-L1 scores and progression free survival for patients receiving first line ICI 
df <- survdat[!is.na(survdat$io_1L),]
surv <- survfit(Surv(PFS_1L_time, PFS_1L_status) ~ PDL1_scores, data = df)
summary(coxph(Surv(PFS_1L_time, PFS_1L_status) ~ PDL1_scores, data = df))

p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
#legend.labs=c('<1%', '1-49%','≥50%'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#E7298A","#7B3294","#885000","#88CCEE","#882255","#CC6677","#7B3254"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS First line IO")
B_1L <- p1

# Association between TME class and progression free survival for patients receiving second line ICI 

df <- survdat[!is.na(survdat$io_2L),]
surv <- survfit(Surv(PFS_2L_time, PFS_2L_status) ~ cluster, data = df)
summary(coxph(Surv(PFS_2L_time, PFS_2L_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS Second line IO")
A_2L <- p1


# Association between PD-L1 scores and progression free survival for patients receiving second line ICI 
df <- survdat[!is.na(survdat$io_2L),]
surv <- survfit(Surv(PFS_2L_time, PFS_2L_status) ~ PDL1_scores, data = df)
summary(coxph(Surv(PFS_2L_time, PFS_2L_status) ~ PDL1_scores, data = df))



p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
#legend.labs=c('<1%', '1-49%','≥50%'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#E7298A","#7B3294","#885000","#88CCEE","#882255","#CC6677","#7B3254"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS Second line IO")
B_2L <- p1

# Association between TME class and progression free survival for patients receiving subsequent line ICI 
df <- survdat[!is.na(survdat$io_SL),]
surv <- survfit(Surv(PFS_SL_time, PFS_SL_status) ~ cluster, data = df)
summary(coxph(Surv(PFS_SL_time, PFS_SL_status) ~ cluster, data = df))


p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
legend.labs=c('Immune Desert', 'Immune Inflamed'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#88CCEE","#882255","#CC6677", "#7B3294","#7B3254","#885000"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS Subsequentline IO")
A_SL <- p1


# Association between PD-L1 scores and progression free survival for patients receiving subsequent line ICI 
df <- survdat[!is.na(survdat$io_SL),]
surv <- survfit(Surv(PFS_SL_time, PFS_SL_status) ~ PDL1_scores, data = df)
summary(coxph(Surv(PFS_SL_time, PFS_SL_status) ~ PDL1_scores, data = df))



p1 <- ggsurvplot(surv, data = df,
legend.title = "",
legend = c(0.73, 0.9),
#legend.labs=c('<1%', '1-49%','≥50%'),
conf.int = T,
pval = TRUE,
risk.table = TRUE,
tables.height = 0.2,
tables.theme = theme_cleantable(),
risk.table.y.text = FALSE,
pval.coord = c(0, 0.03),
# Color palettes. Use custom color: c("#E7B800", "#2E9FDF"),
# or brewer color (e.g.: "Dark2"), or ggsci color (e.g.: "jco")
palette = c("#E7298A","#7B3294","#885000","#88CCEE","#882255","#CC6677","#7B3254"),
ggtheme = theme(text = element_text(size = 15)) +theme_bw() # Change ggplot2 theme
) + xlab("Time (days)") + labs(title = "PFS Subsequentline IO")
B_SL <- p1


# combine all K-M plots into one figure
fig1 <- ggarrange(A_11L$plot,A_2L$plot,A_SL$plot, A_11L$table,A_2L$table, A_SL$table, ncol = 3, nrow = 2,labels = c("A","B","C","","",""),heights = c(3,1))
fig2 <- ggarrange(B_1L$plot,B_2L$plot,B_SL$plot, B_1L$table,B_2L$table, B_SL$table, ncol = 3, nrow = 2,labels = c("D","E","F","","",""),heights = c(3,1))
fig_unsupervised <- ggarrange(fig1,fig2, ncol = 1, nrow = 2, labels = c("",""))
print(fig_unsupervised)
#ggsave(plot = fig_unsupervised, filename = paste(OUTPUT_DIR,"PFS_Unsupervised.pdf",sep = "/"),width = 16, height = 12)
```
# Associations between TME class, PD-L1 scores and treatment response
```{r}
idx = !is.na(survdat$overall_resp) & !is.na(survdat$PDL1_scores)
predicted_combined <- c("R","NR")[as.numeric(as.character(survdat$response_group) == "PD-L1 < 50%, Immune Desert")+1][idx]
predicted_PDL1 <- c("NR","R")[as.numeric(as.character(survdat$PDL1_scores) == "≥50%")+1][idx]
predicted_TME <- c("R","NR")[as.numeric(as.character(survdat$cluster) == "1")+1][idx]
actual <- c("R","NR")[as.numeric(survdat$overall_resp == "NonResponder")+1][idx]
precs <- c()
recs <- c()
f1s <- c()
predicted <- predicted_TME
predicted <- factor(predicted, levels = c("R","NR"))
actual <- factor(actual, levels = c("R","NR"))
result <- confusionMatrix(predicted, actual)
precision <- result$byClass['Pos Pred Value']    
recall <- result$byClass['Sensitivity']
F1 <- c(2*(precision*recall)/(precision+recall))
names(F1) <- "F1"
precs <- c(precs, precision)
recs <- c(recs, recall)
f1s <- c(f1s, F1)
rval <- epi.tests(result$table, conf.level = 0.95)
print(rval)



precs <- c()
recs <- c()
f1s <- c()
predicted <- predicted_PDL1
predicted <- factor(predicted, levels = c("R","NR"))
actual <- factor(actual, levels = c("R","NR"))
result <- confusionMatrix(predicted, actual)
precision <- result$byClass['Pos Pred Value']    
recall <- result$byClass['Sensitivity']
F1 <- c(2*(precision*recall)/(precision+recall))
names(F1) <- "F1"
precs <- c(precs, precision)
recs <- c(recs, recall)
f1s <- c(f1s, F1)
rval <- epi.tests(result$table, conf.level = 0.95)
print(rval)


precs <- c()
recs <- c()
f1s <- c()
predicted <- predicted_combined
predicted <- factor(predicted, levels = c("R","NR"))
actual <- factor(actual, levels = c("R","NR"))
result <- confusionMatrix(predicted, actual)
precision <- result$byClass['Pos Pred Value']    
recall <- result$byClass['Sensitivity']
F1 <- c(2*(precision*recall)/(precision+recall))
names(F1) <- "F1"
precs <- c(precs, precision)
recs <- c(recs, recall)
f1s <- c(f1s, F1)
rval <- epi.tests(result$table, conf.level = 0.95)
print(rval)
```

